---
title: "Linear_Models"
author: "Maliha Safdar"
date: "2025-11-06"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(modelr)
set.seed(1)
```

## Model fitting 

The code below loads and cleans the Airbnb data, which we’ll use as a primary example for fitting linear models.

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type)
```


A good place to start is to consider price as an outcome that may depend on rating and borough. We fit that initial model in the following code.

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

The lm function begins with the formula specification – outcome on the left of the ~ and predictors separated by + on the right. As we’ll see shortly, interactions between variables can be specified using *. You can also specify an intercept-only model (outcome ~ 1), a model with no intercept (outcome ~ 0 + ...), and a model using all available predictors (outcome ~ .).

R will treat categorical (factor) covariates appropriately and predictably: indicator variables are created for each non-reference category and included in your model, and the factor level is treated as the reference. As with ggplot, being careful with factors is therefore critical!

```{r}
nyc_airbnb = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough), ## fct_infreq says putting categories in orders of how frequently they appear in the data for example if Manhattan appears the most it is #1 on the factor, then if Brooklyn is the second most that appears it is given #2 and so on.
    room_type = fct_infreq(room_type))

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

It’s important to note that changing reference categories won’t change “fit” or statistical sigificance, but can affect ease of interpretation.

## Tidying output
The output of a lm is an object of class lm – a very specific list that isn’t a dataframe but that can be manipulated using other functions. Some common functions for interacting with lm fits are below, although we omit the output.

```{r, eval = FALSE}
summary(fit)
names(summary(fit))
summary(fit)$coef
coef(fit)
fitted.values(fit)
summary(fit)[["df"]]
```

The reason that we omit the output is that it’s a huge pain to deal with. summary produces an object of class summary.lm, which is also a list – that’s how we extracted the coefficients using summary(fit)$coef. coef produces a vector of coefficient values, and fitted.values is a vector of fitted values. None of this is tidy.

It’s helpful to know about the products of lm and to know there are a range of ways to interact with models in base R. That said, for the most part it’s easiest to use tidy tools.

The broom package has functions for obtaining a quick summary of the model and for cleaning up the coefficient table.

```{r}
fit |>
  broom :: tidy() |>
  mutate(
    term = str_replace(term, "borough", "Bourough:")
  ) |>
  select(term, estimate, p.value) |>
  knitr::kable(digits = 3)

fit |>
  broom :: glance()

```

## Diagnostics

```{r}

modelr::add_residuals(nyc_airbnb,fit) |> 
modelr::add_predictions(nyc_airbnb, fit) |>
  filter(resid <1000) |>
  ggplot(aes(x = resid)) +
  geom_histogram()


modelr::add_residuals(nyc_airbnb, fit) |> 
modelr::add_predictions(nyc_airbnb, fit) |>
  filter(resid <1000) |>
  ggplot(aes(x = borough, y = resid)) +
  geom_violin()


modelr::add_residuals(nyc_airbnb, fit) |> 
modelr::add_predictions(nyc_airbnb, fit) |>
  filter(resid <1000) |>
  ggplot(aes(x = starts, y = resid)) +
  geom_point()

```

## Hypothesis Testing


We’ll comment briefly on hypothesis testing. Model summaries include results of t-tests for single coefficients, and are the standard way of assessing statistical significance.

Testing multiple coefficients is somewhat more complicated. A useful approach is to use nested models, meaning that the terms in a simple “null” model are a subset of the terms in a more complex “alternative” model. The are formal tests for comparing the null and alternative models, even when several coefficients are added in the alternative model. Tests of this kind are required to assess the significance of a categorical predictor with more than two levels, as in the example below.

```{r}
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)

anova(fit_null, fit_alt) |>
  broom :: tidy()
```

## Interactions vs nested data


```{r}
fit_interactions = 
  lm(price ~ stars * borough + room_type * borough, data = nyc_airbnb)

fit_interactions |>
  broom :: tidy()
```


Let's look at Brooklyn first
```{r}
nyc_airbnb |>
  filter(borough == "Brooklyn") |>
  lm(price ~ stars + room_type, data = _) |> #"_" for data, the first dataframe, tells the pipe that _ is a place holder 
  broom :: tidy() #the output is just for Brooklyn

nyc_airbnb |>
  filter(borough == "Manhattan") |>
  lm(price ~ stars + room_type, data = _) |> #"_" for data, the first dataframe, tells the pipe that _ is a place holder 
  broom :: tidy() #the output is just for Manhattan


nyc_airbnb |>
  filter(borough == "Bronx") |>
  lm(price ~ stars + room_type, data = _) |> #"_" for data, the first dataframe, tells the pipe that _ is a place holder 
  broom :: tidy() #the output is just for Bronx

nyc_airbnb |>
  filter(borough == "Queens") |>
  lm(price ~ stars + room_type, data = _) |> #"_" for data, the first dataframe, tells the pipe that _ is a place holder 
  broom :: tidy() #the output is just for Queens


#We want something that will not have us copy and paste the code for each borough
```

Write a short function

```{r}
lm_airbnb = function(df) {
  lm(price ~ stars + room_type, data = df)
}

nyc_airbnb |>
  filter(borough == "Queens") |>
  lm_airbnb() |>
  broom ::tidy()
```



Create a list of dataframes, and iterate to fit the model each time.

```{r}
nested_lm_results = nyc_airbnb |>
  nest(data = - borough) |>
  mutate(
    fits = map(data, lm_airbnb),
    results = map(fits, broom :: tidy)
  ) |>
  select(borough, results) |>
  unnest(results)

```

DO some untidying

```{r}
nested_lm_results |>
  select(borough, term, estimate) |>
  pivot_wider(
    names_from = term,
    values_from = estimate
  )
```

Use an *anonymous* function instead of `lm_airbnb`


```{r}
nested_lm_results = nyc_airbnb |>
  nest(data = - borough) |>
  mutate(
    fits = map(data, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(fits, broom :: tidy)
  ) |>
  select(borough, results) |>
  unnest(results)

```


Let's do an even more extreme example

```{r}
manhattan_analysis =
  nyc_airbnb |>
  filter(
    borough == "Manhattan"
  ) |>
  nest(data = -neighborhood) |>
  mutate(
    fits = map(data, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(fits, broom :: tidy)
  ) |>
  select(neighborhood, results) |>
  unnest(results)
```


Make a plot:

```{r}
manhattan_analysis |>
  filter(term == "stars") |>
  mutate(
    neighborhood = fct_reorder(neighborhood, estimate)
  ) |>
  ggplot(aes(x = neighborhood, y = estimate)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

